## **R Libraries: ggplot2**

The R library `ggplot2` is a super useful tool we can use to create visualizations and explore our data.

```{bash}
install.packages("tidyverse")
library("tidyverse")
```

More things we can do to get more information on R and what we're doing:

```{bash}
help()
```

We'll need a few more packages for our exploration usage, too. We can install them all at once by using the `c()` function, which creates a list. We'll open these packages with the `library()` function later.

```{bash}
install.packages(c("nycflights13", "gapminder", "Lahman"))
```

Built in to our tidyverse are several data sets that we can play around with. In the below code block, we create a variable `myplot` and save the `mpg` database to it.

```{r}
myplot <- mpg
myplot
```

**Something you should be aware of and have open (perhaps on a second monitor, or you can print them out) is the ggplot2 cheat sheet!** <https://github.com/rstudio/cheatsheets/blob/main/data-visualization-2.1.pdf>

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))
View(myplot)
```

```{r}
mtcars 

```

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, color = displ < 5))
#also try color = class
```

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = rgb(0.7,1,0.3))) + 
  facet_wrap(~ class, nrow = 2)
#try shape = class, color = class
```

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = drv)) + 
  facet_grid(drv ~ .)
```

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, linetype = drv)) + 
  geom_point(mapping = aes(color = class)) +
  geom_smooth(se = TRUE)
```

```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut))
```

```{r}

```

Whoa, where'd count come from? **Count is not a variable or attribute of the variable diamonds!** Count comes automatically with a lot of graphs as an attribute. You can verify what I'm saying with `?diamonds` in the console. It shows up because it is a discrete data set, rather than a continuous data set that we have been using.

Any attribute of a data set that is algorithmically calculated is called a *stat*, which is short for a statistical transformation. Many of the **geom** functions have stats built in, and many stats display geoms. For example, the above code block used a geom, but this one uses a stat, and results in an identical chart:

```{r}
ggplot(data = diamonds) + 
  stat_count(mapping = aes(x = cut))
```

Let's add color here!

```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x= cut, fill = cut))
```

What if we add to a different variable than cut?

```{r}
pie = ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity), 
           position = 'fill'
  ) + 
  theme(aspect.ratio = 1) + 
  labs(x = NULL, y = NULL)

pie + coord_polar()
```

**For your consideration...** Is the above a good graph?

```{r}
alt- becomes <- (option- on mac)
```

## Homework

```{r}
# Find a builtin data set for students to explore; use their cheat sheet from stats to label continuous and discrete data; Analyze that data set, create graphs, find a hypothesis. Use graphs to tell me things about 3 data sets. 

View(nycflights13::flights)
```
